---
title: "Background"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{background}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#library(gediamazonia)
library(sf)
library(ggplot2)
```

## Overview

The goal of this project is to collect and analyze GEDI data within an area of 
interest in and around the Cordillera Azul National Park in Peru.
```{r, fig.cap="Cordillera Azul NP (blue), park buffer (lightblue), cacao plantation zone (red outline), and area of interest (purple outline)."}
data(studyarea)
dims <- do.call(rbind, lapply(studyarea, sf::st_bbox))
limits <- lapply(list(c(1, 3), c(2, 4)), function(x) range(dims[, x]))

ggplot() + geom_sf(data = studyarea$np, fill = "blue") + 
  geom_sf(data = studyarea$buffer, fill = "lightblue") + 
  geom_sf(data = studyarea$cacao_zone, color = "red", fill = "transparent") +
  geom_sf(data = studyarea$aoi, color = "purple", fill = "transparent")
```

GEDI should be collected along a West to East Transect that falls within the 
area of interest (AOI) and includes the park, buffer and cacao farming zone. 

## Study Area
Cordillera Azul National Park in Peru was established in 1963. It spans 
2,718,00 ha and protects the Ucayali moist forests ecoregion. The national 
park primarily consists of montane forests, but also has sections of dry forest 
to the West and Amazon rainforest to the East. 

## GEDI
GEDI is a full-waveform lidar instrument which launched on December 5th, 2018. 
It currently is the highest resolution and densest sampling of any lidar ever 
put into orbit. It has a horizontal accuracy within +/- 9m, and minimizes 
off-nadir pointing to less than 6 degrees to avoid inaccuracy. From GEDI 
waveforms, four types of vegetation structure information can be extracted: 
surface topography, canopy height metrics, canopy cover metrics, and vertical 
structure metrics. The raw waveforms and derived products are available online. 
These products include Geolocated waveforms (L1B), Elevation and Height Metrics 
(L2A), Canopy Cover Fraction and Leaf Area Index (L2B), and footprint-level 
above-ground biomass (L4A). Each of these is available at 25 meter diameter 
resolution. There are also grided forms at 1 km resolution.



