---
title: "Visualizing GEDI Data"
author: "Amanda Payton, Sam Watson, Wenqu Chen"
date: "May 2, 2020"
output: html_document
vignette: >
  %\VignetteIndexEntry{Visualizing GEDI Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Overview

### This vignette uses the clipped shapefiles from the GEDI data vignette and 
### visualizes the data through maps and box plots.

# Setup

## Package Installation and library loading

```{r}
#install.packages("rGEDI")
library(rGEDI)
#install.packages("geospaar")
library(geospaar)
library(ggplot2)
```

## Set Folder paths

```{r}
#check working directory
getwd() 
#If needed, set your working directory to the gediamazonia project folder

data_fold = paste0(getwd(), "/data")
working_fold = paste0(getwd(), "/working")
```

## Read in GEDI Shapefiles and Study Area Data

```{r}
data("studyarea")
data("roads")
data('gedi_ss')

#Data clipped to AOI
level1b <- st_read(dsn = paste0(data_fold, "/aoi_GEDI_1B.shp"))
head(level1b)

level2a <- st_read(dsn = paste0(data_fold, "/aoi_GEDI_2A.shp"))
head(level2a)

level2b <- st_read(dsn = paste0(data_fold, "/aoi_GEDI_2B.shp"))
head(level2b)

level2bPAI <- st_read(dsn = paste0(data_fold, "/aoi_GEDI_2B_PAI.shp"))
head(level2bPAI)

level2bPAVD <- st_read(dsn = "/aoi_GEDI_2B_PAVD.shp")
head(level2bPAVD)

#Data clipped to entire study region
level2a_SA <- st_read(dsn = paste0(data_fold, "/SA_GEDI_2A.shp"))
head(level2a_SA)

level2b_SA <- st_read(dsn = paste0(data_fold, "/SA_GEDI_2B.shp"))
head(level2b_SA)
```

# Visualizations

## Map Study Area

```{r}
#legtext <- c("National Park", "Park Buffer", "Cacao Zone")
#legfill <- c("dark green", "grey90", "tan")

#Create map of Study Area
p <- ggplot() + geom_sf(data = studyarea$np, fill = "dark green") + 
  geom_sf(data = studyarea$buffer, color = "grey", fill = "light blue") + 
  geom_sf(data = studyarea$cacao_zone, color = "brown", 
          fill = alpha("tan", .2)) +   
  geom_sf(data = roads, color = "black") +
  theme_minimal() +
  ggtitle("Study Region: Cordillera Azul \n National Park") + 
     theme(plot.title = element_text(face = "bold"))
  
#Export to Working Folder
ggsave(p, filename = paste0(working_fold, "/study_area.png"), height = 4,
       width = 5, dpi = 300)
```

## Map Variables

```{r}
ggplot() + geom_sf(data = studyarea$np, color = "grey", fill = "grey70") + 
  geom_sf(data = studyarea$buffer, color = "grey", fill = "grey80") + 
  geom_sf(data = studyarea$cacao_zone, color = "grey", fill = "grey90") +   
  geom_sf(data = studyarea$aoi, color = "grey") + 
  geom_sf(data = roads, color = "grey") +
  theme_minimal() +
  #geom_sf(data = level1b, aes(color = elvtn_0)) #level 1B Elevation data
  #geom_sf(data = level2a, aes(color = rh25)) #level 2A relative height 25
  #geom_sf(data = level2a, aes(color = rh50)) #level 2A relative height 50
  #geom_sf(data = level2a, aes(color = rh75)) #level 2A relative height 75
  #geom_sf(data = level2a, aes(color = rh100)) #level 2A relative height 100
  geom_sf(data = level2b, aes(color = pai)) #level 2B plant area index


```

## Create Boxplots - Relative Height and Zones

### Preprocessing

```{r}
#Intersect level 2A data shapefile with zones 
cacao_intersect <- st_intersection(x = studyarea$cacao_zone, y = level2a_SA)
park_intersect <- st_intersection(x = studyarea$np, y = level2a_SA)
buffer_intersect <- st_intersection(x = studyarea$buffer, y = level2a_SA)

#Create Relative Height 25 Variable
cacao_RH25 <- cacao_intersect$rh25 %>% as_data_frame(.) %>% 
  mutate(zone = "Cacao", level = "rh25")
park_RH25 <- park_intersect$rh25 %>% as_data_frame(.) %>% 
  mutate(zone = "Park", level = "rh25")
buffer_RH25 <- buffer_intersect$rh25 %>% as_data_frame(.) %>% 
  mutate(zone = "Buffer", level = "rh25")

#Create Relative Height 50 Variable
cacao_RH50 <- cacao_intersect$rh50 %>% as_data_frame(.) %>% 
  mutate(zone = "Cacao", level = "rh50")
park_RH50 <- park_intersect$rh50 %>% as_data_frame(.) %>% 
  mutate(zone = "Park", level = "rh50")
buffer_RH50 <- buffer_intersect$rh50 %>% as_data_frame(.) %>% 
  mutate(zone = "Buffer", level = "rh50")

#Create Relative Height 75 Variable
cacao_RH75 <- cacao_intersect$rh75 %>% as_data_frame(.) %>% 
  mutate(zone = "Cacao", level = "rh75")
park_RH75 <- park_intersect$rh75 %>% as_data_frame(.) %>% 
  mutate(zone = "Park", level = "rh75")
buffer_RH75 <- buffer_intersect$rh75 %>% as_data_frame(.) %>% 
  mutate(zone = "Buffer", level = "rh75")

#Create Relative Height 100 Variable
cacao_RH100 <- cacao_intersect$rh100 %>% as_data_frame(.) %>% 
  mutate(zone = "Cacao", level = "rh100")
park_RH100 <- park_intersect$rh100 %>% as_data_frame(.) %>% 
  mutate(zone = "Park", level = "rh100")
buffer_RH100 <- buffer_intersect$rh100 %>% as_data_frame(.) %>% 
  mutate(zone = "Buffer", level = "rh100")

#bind all RH levels
rh_bound <- rbind(cacao_RH25, park_RH25, buffer_RH25, cacao_RH50, 
                  park_RH50, buffer_RH50, cacao_RH75, park_RH75, 
                  buffer_RH75, cacao_RH100, park_RH100, buffer_RH100)
```

### Plot

```{r}
#plot RH levels
rhplot <- rh_bound %>% 
  ggplot() + geom_boxplot(aes(x = level, y = value, fill = zone)) + 
  xlab(NULL) + scale_fill_manual(values = c("light blue", "tan", "dark green")) + 
  ggtitle("Relative Height Metrics Between Geometries")

#save plot
ggsave("rhplot.png", rhplot, width = 4, height = 2)
```

## Create Boxplots - Plant Area Index and Zones

<!-- ```{r} -->
<!-- level2BVPM_clip_gb <- clipLevel2BVPMGeometry(level2BVPM, polygon_spdf, split_by = split_by) -->

<!-- #Remove -9999 from PAI metric and NAs from lat and long -->
<!-- PAI_NA <- level2BVPM_clip_gb[level2BVPM_clip_gb$pai == -9999] <- NA -->
<!-- level2BVPM_clip_gb$longitude_bin0[is.na(level2AM_clip_gb$longitude_bin0)] <- 0 -->
<!-- level2BVPM_clip_gb$latitude_bin0[is.na(level2AM_clip_gb$latitude_bin0)] <- 0 -->
<!-- level2BVPM_clip_gb$shot_number <- paste0(level2AM_clip_gb$shot_number) -->

<!-- level2BVPM_spdf <- SpatialPointsDataFrame(cbind(level2AM_clip_gb$longitude_bin0, level2AM_clip_gb$latitude_bin0), data=level2AM_clip_gb) -->
<!-- raster::shapefile(level2BVPM_spdf,paste0(outdir,"\\level2bvpm_Clipped")) -->

<!-- #read in clipped shapefile and project -->
<!-- level2BVPMClipped_shape<- st_read(paste0(outdir,"/level2BVPM_Clipped.shp")) -->
<!-- reference <- st_crs(cacao_zone) -->
<!-- st_crs(level2BVPMClipped_shape) <- reference -->

<!-- ##PAI PRE - PROCESSING  -->
<!-- #clip clipped shapefile to different zones -->
<!-- cacao_PAI <- st_intersection(x = cacao_zone, y = level2BVPMClipped_shape) -->
<!-- park_PAI <- st_intersection(x = park, y = level2BVPMClipped_shape) -->
<!-- buffer_PAI <- st_intersection(x = buffer, y = level2BVPMClipped_shape) -->

<!-- #make PAI dataframes for each zone -->
<!-- cacao_PAI <- cacao_PAI$pai %>% as_data_frame(.) %>% mutate(zone = "Cacao") -->
<!-- park_PAI <- park_PAI$pai %>% as_data_frame(.) %>% mutate(zone = "Park") -->
<!-- buffer_PAI <- buffer_PAI$pai %>% as_data_frame(.) %>% mutate(zone = "Buffer") -->
<!-- rbind(cacao_PAI, park_PAI, buffer_PAI) -->

<!-- #set color pallette -->
<!-- clrs <- c("light blue", "tan", "dark green") -->

<!-- #bind rows and plot, then save offline -->
<!-- rbind(cacao_PAI, park_PAI, buffer_PAI) %>% ggplot() + geom_boxplot(aes(x = zone, y = value, fill = zone))+ xlab(NULL)+  ggtitle("Plant Area Index Metrics by Zone") + scale_fill_manual(values = clrs) -->
<!-- ggsave("pai_zones.png", pai_plot, width = 3, height = 2) -->

<!-- ``` -->

## Create Boxplots - Relative Height and Road Distance -->

<!-- ```{r} -->
<!-- #Create dataframes for each rh level -->
<!-- rh25 <- gedi_ss$rh25 %>% as_data_frame(.) %>% mutate(level = "rh25") -->
<!-- rh50 <- gedi_ss$rh50 %>% as_data_frame(.) %>% mutate(level = "rh50") -->
<!-- rh75 <- gedi_ss$rh75 %>% as_data_frame(.) %>% mutate(level = "rh75") -->
<!-- rh100 <- gedi_ss$rh100 %>% as_data_frame(.) %>% mutate(level = "rh100") -->

<!-- #bind all of the rh levels -->
<!-- allrhlevels <- rbind(rh25, rh50, rh75, rh100) -->
<!-- gediss_bind <- cbind(distances, allrhlevels) -->
<!-- road_rh_levels <- gediss_bind[-c(1)] -->
<!-- roads_rhplot <- road_rh_levels %>%  -->
<!--   ggplot() + geom_boxplot(aes(x = level, y = value, fill = distances)) +  -->
<!--   xlab(NULL) + scale_fill_manual(values = c("#f0f9e8", "#ccebc5", "#a8ddb5", "#7bccc4", "#43a2ca", "#0868ac")) + ggtitle("Relative Height MetricsDist from Roads") + labs(fill = "Distance Rank") -->
<!-- ggsave(paste0(getwd(),"/external/presentation/roads_rhplot.png"), roads_rhplot, width = 5, height = 3) -->
<!-- ``` -->

