---
title: "Getting and preprocessing GEDI data"
author: "Amanda Payton, Sam Watson, Wenqu Chen"
date: "May 2, 2020"
output: html_document
vignette: >
  %\VignetteIndexEntry{Getting and preprocessing GEDI data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Overview

### This vignette covers how to download GEDI data, complete pre-processing, 
### convert hd5 files into shapefile, and clip the shapefile using the 
### Cordillera Azul National Park as the area of interest. This code is based 
### off of the rGEDI tutorial created by Carlos Alberto Silva.

# Setup

## Package Installation and library loading

```{r}
#install.packages("rGEDI")
library(rGEDI)
#install.packages("geospaar")
library(geospaar)
```

## Set Folder paths

```{r}
#check working directory
getwd() 
#If needed, set your working directory to the gediamazonia project folder

raw_data_fold = paste0(getwd(), "/data-raw")
data_fold = paste0(getwd(), "/data")
```

## Set Area of Interest and Date Range

```{r}
#load data
data("studyarea")

#get bounding box coordinates
st_bbox(studyarea$aoi)

#set coordinates to match bounding box
y_max <- -7.999644
y_min <- -9.437084
x_min <- -76.157869 
x_max <- -74.099156  

#pick date range
daterange = c("2019-08-01", "2019-08-15")
```

# Download GEDI Data

You will be asked to enter your USGS Earth Explorer login information when you 
download GEDI data in this project for the first time.

## Level 1B Geolocated Waveforms

```{r}
gLevel1B <- gedifinder(product = "GEDI01_B", y_max, x_max, y_min, x_min, 
                       version = "001", daterange = daterange)

#gediDownload(filepath = gLevel1B, outdir = raw_data_fold)
```

## Level 2A Elevation Height Metrics

```{r}
gLevel2A <- gedifinder(product = "GEDI02_A", y_max, x_max, y_min, x_min, 
                       version = "001", daterange = daterange)

#gediDownload(filepath = gLevel2A, outdir = raw_data_fold)
```

## Level 2B Canopy Cover and Vertical Profile Metrics

```{r}
gLevel2B <- gedifinder(product = "GEDI02_B", y_max, x_max, y_min, x_min, 
                       version = "001", daterange = daterange)

#gediDownload(filepath = gLevel2B, outdir = raw_data_fold)
```

# Read in Dowloaded Data and Display Head

## Level 1B

```{r}
#Read in GEDI data
gedilevel1b <- readLevel1B(level1Bpath = file.path(
  raw_data_fold, "GEDI01_B_2019216034003_O03638_T04321_02_003_01.h5")) 
  #if you altered the date rage or aoi, the file name will change

#Get Full-Waveform Geolocation
level1bGeo <- getLevel1BGeo(level1b = gedilevel1b, select = c("elevation_bin0"))

#Display Head
head(level1bGeo)
```

## Level 2A

```{r}
#Read in GEDI data
gedilevel2a <- readLevel2A(level2Apath = file.path(
  raw_data_fold, "GEDI02_A_2019216034003_O03638_T04321_02_001_01.h5"))
  #if you altered the date rage or aoi, the file name will change

#Get Elevation and Height Metrics
level2aGeo <- getLevel2AM(gedilevel2a)

#Display Head
head(level2aGeo[,c("beam","shot_number","elev_highestreturn",
                   "elev_lowestmode","rh100")])
```

## Level 2B

```{r}
#Read in GEDI data
gedilevel2b <- readLevel2B(level2Bpath = file.path(
  raw_data_fold, "GEDI02_B_2019216034003_O03638_T04321_02_001_01.h5"))
  #if you altered the date rage or aoi, the file name will change

#Get Canopy Cover and Vertical Profile Metrics
level2BGeo <- getLevel2BVPM(gedilevel2b)
head(level2BGeo[,c("beam","shot_number","pai","fhd_normal","omega",
                   "pgap_theta","cover")])

```

# Convert Data into a Spatial Data Frame and Export it as a Shapefile

## Level 1B 

```{r}
#Convert shot_number as "integer64" to "character"
level1bGeo$shot_number <- paste0(level1bGeo$shot_number)

#Convert NA values in coordinates to 0
level1bGeo$longitude_bin0[is.na(level1bGeo$longitude_bin0)] <- 0
level1bGeo$latitude_bin0[is.na(level1bGeo$latitude_bin0)] <- 0

#Convert level1bGeo as data.table to SpatialPointsDataFrame
level1bGeo_spdf <- SpatialPointsDataFrame(cbind(level1bGeo$longitude_bin0, 
                                                level1bGeo$latitude_bin0), 
                                          data = level1bGeo)
#Export level1bGeo as ESRI Shapefile
raster::shapefile(level1bGeo_spdf,paste0(data_fold,"/GEDI01_B"))
```

## Level 2A

```{r}
#Convert shot_number as "integer64" to "character"
level2aGeo$shot_number <- paste0(level2aGeo$shot_number)
level2aGeo$shot_number <- paste0(level2aGeo$shot_number)

#Convert NA values in coordinates to 0
level2aGeo$lon_lowestmode[is.na(level2aGeo$lon_lowestmode)] <- 0
level2aGeo$lat_lowestmode[is.na(level2aGeo$lat_lowestmode)] <- 0

#Convert Elevation and Height Metrics as data.table to SpatialPointsDataFrame
level2A_spdf <- SpatialPointsDataFrame(cbind(level2aGeo_1$lon_lowestmode, 
                                              level2aGeo_1$lat_lowestmode), 
                                        data = level2aGeo_1)

#Export Elevation and Height Metrics as ESRI Shapefile
raster::shapefile(level2A_spdf, paste0(data_fold, "/GEDI02_A"))

```

## Level 2B

```{r}
#Convert shot_number as "integer64" to "character"
level2BGeo$shot_number <- paste0(level2BGeo$shot_number)

#Convert NA values in coordinates to 0
level2BGeo$longitude_lastbin[is.na(level2BGeo$longitude_lastbin)] <- 0
level2BGeo$latitude_lastbin[is.na(level2BGeo$latitude_lastbin)] <- 0

#Convert GEDI Vegetation Profile Biophysical Variables as data.table to 
#SpatialPointsDataFrame
level2B_spdf <- SpatialPointsDataFrame(cbind(level2BGeo$longitude_lastbin, 
                                             level2BGeo$latitude_lastbin), 
                                       data = level2BGeo)

# Export GEDI Vegetation Profile Biophysical Variables as ESRI Shapefile
raster::shapefile(level2B_spdf,paste0(data_fold,"/GEDI02_B"))
```

## Level 2B: Get Plant Area Index and Plant Area Volume Density

```{r}
#Plant Area Index
level2BPAIProfile <- getLevel2BPAIProfile(gedilevel2b)
head(level2BPAIProfile[, c("beam", "shot_number", "pai_z0_5m", 
                           "pai_z5_10m")])

#Plant Area Volume Density
level2BPAVDProfile <- getLevel2BPAVDProfile(gedilevel2b)
head(level2BPAVDProfile[, c("beam", "shot_number", "pavd_z0_5m", 
                            "pavd_z5_10m")])

#Convert shot_number as "integer64" to "character"
level2BPAIProfile$shot_number <- paste0(level2BPAIProfile$shot_number)
level2BPAVDProfile$shot_number <- paste0(level2BPAVDProfile$shot_number)

#Convert NA values in coordinates to 0
level2BPAVDProfile$lat_lowestmode[is.na(level2BPAVDProfile$lat_lowestmode)] <- 0
level2BPAVDProfile$lon_lowestmode[is.na(level2BPAVDProfile$lon_lowestmode)] <- 0

level2BPAIProfile$lat_lowestmode[is.na(level2BPAIProfile$lat_lowestmode)] <- 0
level2BPAIProfile$lon_lowestmode[is.na(level2BPAIProfile$lon_lowestmode)] <- 0

#Convert PAI and PAVD Profiles as data.table to SpatialPointsDataFrame
level2BPAIProfile_spdf <- SpatialPointsDataFrame(cbind(
  level2BPAIProfile$lon_lowestmode, level2BPAIProfile$lat_lowestmode), 
  data = level2BPAIProfile)
level2BPAVDProfile_spdf <- SpatialPointsDataFrame(cbind(
  level2BPAVDProfile$lon_lowestmode, level2BPAVDProfile$lat_lowestmode), 
  data = level2BPAVDProfile)

#Export PAI and PAVD Profiles as ESRI Shapefile.
raster::shapefile(level2BPAIProfile_spdf, paste0(
  data_fold, "/GEDI02_B_PAIProfile"))
raster::shapefile(level2BPAVDProfile_spdf, paste0(
  data_fold, "/GEDI02_B_PAVDProfile"))
```

# Clip Shapefiles to the AOI

## Level 1B

```{r}
#read in shapefile
level1bShp <- st_read(dsn = paste0(data_fold, "/GEDI01_B.shp"))

#get AOI coordinate system and set shapefile to that 
reference <- st_crs(studyarea$aoi)
st_crs(level1bShp) <- reference

#make a polygon with AOI coordinates
coords <- cbind("x" = c(x_max, x_min, x_min, x_max, x_max), 
                "y" = c(y_max, y_max, y_min, y_min, y_max))
aoi_bbox <- st_polygon(x = list(coords)) %>% st_sfc %>% st_sf(ID = 1, 
                                                              crs = 4326)

#clip shapefile to AOI polygon
aoi_1bLidar <- st_intersection(x = aoi_bbox, y = level1bShp)

#export clipped shapefile
st_write(obj = aoi_1bLidar, dsn = file.path(data_fold, "AOI_GEDI_1B.shp"))
```

## Level 2A

```{r}
#read in shapefile
level2aShp <- st_read(dsn = paste0(data_fold, "/GEDI02_A1.shp"))

#get AOI coordinate system and set shapefile to that 
reference <- st_crs(studyarea$aoi)
st_crs(level2aShp) <- reference

#make a polygon with AOI coordinates
coords <- cbind("x" = c(x_max, x_min, x_min, x_max, x_max), 
                "y" = c(y_max, y_max, y_min, y_min, y_max))
aoi_bbox <- st_polygon(x = list(coords)) %>% st_sfc %>% st_sf(ID = 1, 
                                                              crs = 4326)

#clip shapefile to AOI polygon
aoi_2aLidar <- st_intersection(x = aoi_bbox, y = level2aShp)

#export clipped shapefile
st_write(obj = aoi_2aLidar, dsn = file.path(data_fold,"aoi_GEDI_2A.shp"))
```

## Level 2B, PAI, and PAVD

```{r}
#read in shapefile
level2bPAVDshp <- st_read(dsn = paste0(data_fold, "/GEDI02_B_PAVDProfile.shp"))
level2bPAIShp <- st_read(dsn =  paste0(data_fold, "/GEDI02_B_PAIProfile.shp"))
level2bShp <- st_read(dsn = paste0(data_fold, "/GEDI02_B.shp"))

#get AOI coordinate system and set shapefile to that 
reference <- st_crs(studyarea$aoi)
st_crs(level2bPAVDshp) <- reference
st_crs(level2bPAIShp) <- reference
st_crs(level2bShp) <- reference

#make a polygon with AOI coordinates
coords <- cbind("x" = c(x_max, x_min, x_min, x_max, x_max), 
                "y" = c(y_max, y_max, y_min, y_min, y_max))
aoi_bbox <- st_polygon(x = list(coords)) %>% st_sfc %>% st_sf(ID = 1, 
                                                              crs = 4326)

#clip shapefile to AOI polygon
aoi_2bPAVD <- st_intersection(x = aoi_bbox, y = level2bPAVDshp)
aoi_2bPAI <- st_intersection(x = aoi_bbox, y = level2bPAIShp)
aoi_2b <- st_intersection(x = aoi_bbox, y = level2bShp)

#export clipped shapefile
st_write(obj = aoi_2bPAVD, dsn = file.path(data_fold, "aoi_GEDI_2B_PAVD.shp"))
st_write(obj = aoi_2bPAI, dsn = file.path(data_fold, "aoi_GEDI_2B_PAI.shp"))
st_write(obj = aoi_2b, dsn = file.path(data_fold, "aoi_GEDI_2B.shp"))
```
