---
title: "accessingGEDI"
author: "Sam Watson"
date: "April 8, 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

load necessary packages
```{r}
library(rGEDI)
library(sf)
library(sp)
```

load AOI data and find GEDI data
```{r}
data("studyarea")
aoi <- studyarea$aoi
st_bbox(aoi)

study_Area <- st_read(dir("F:\\R_Project", pattern = "bb_points.shp", full.names = TRUE))
st_crs(study_Area) <- st_crs(aoi)

#bolunding box coordinates
ul_lat <- -7.999644
lr_lat <- -9.437084
ul_lon <- -74.099156
lr_lon <- -76.157869

#Data Range
daterange=c("2019-07-10", "2019-07-15")
#Get path to GEDI

gLevel1B <- gedifinder(product = "GEDI01_B", ul_lat, ul_lon, lr_lat, lr_lon, version = "001", daterange = daterange)
gLevel2A <- gedifinder(product = "GEDI02_A", ul_lat, ul_lon, lr_lat, lr_lon, version = "001", daterange = daterange)
gLevel2B <- gedifinder(product = "GEDI02_B", ul_lat, ul_lon, lr_lat, lr_lon, version = "001", daterange = daterange)
```

Downloading the data
```{r}
outdir = "F:\\R_Project\\July_2019"
# gediDownload(filepath = gLevel1B, outdir = outdir)
# gediDownload(filepath = gLevel2A, outdir = outdir)
# gediDownload(filepath = gLevel2B, outdir = outdir)
```

reading GEDI data
```{r}
# Reading GEDI data
gedilevel1b <- readLevel1B(level1Bpath = paste0(outdir, "\\GEDI01_B_2019195233920_O03325_T05109_02_003_01.h5"))
gedilevel2a <- readLevel2A(level2Apath = paste0(outdir, "\\GEDI02_A_2019195233920_O03325_T05109_02_001_01.h5"))
gedilevel2b <- readLevel2B(level2Bpath = paste0(outdir, "\\GEDI02_B_2019195233920_O03325_T05109_02_001_01.h5"))
```

GEDI pulse location level 1B
```{r}
#get geogprahic information from gedi data and extracting elevation
level1bGeo <- getLevel1BGeo(level1b = gedilevel1b, select = c("elevation_bin0"))
head(level1bGeo)


# Converting shot_number as "integer64" to "character"
level1bGeo$shot_number <- paste0(level1bGeo$shot_number)

# Converting level1bGeo as data.table to SpatialPointsDataFrame
library(sp)
level1bGeo$longitude_bin0[is.na(level1bGeo$longitude_bin0)] <- 0
level1bGeo$latitude_bin0[is.na(level1bGeo$latitude_bin0)] <- 0

#clipping level 1B GEDI to AOI
level1bGeo_clip_bb <- clipLevel1BGeo(level1bGeo, xmin, xmax, ymin, ymax)

level1bGeo_spdf <- SpatialPointsDataFrame(cbind(level1bGeo$longitude_bin0, level1bGeo$latitude_bin0),
                                        data = level1bGeo)

# Exporting level1bGeo as ESRI Shapefile
#raster::shapefile(level1bGeo_spdf,paste0(outdir, "\\GEDI01_B_2019108080338_O01964_T05337_02_003_01_sub"))

#leaflet visualization of clipped pule locations
library(leaflet)
library(leafsync)

leaflet() %>%
  addCircleMarkers(level1bGeo_clip_bb$longitude_bin0,
                   level1bGeo_clip_bb$latitude_bin0,
                   radius = 1,
                   opacity = 1,
                   color = "red")  %>%
  addScaleBar(options = list(imperial = FALSE)) %>%
  addProviderTiles(providers$Esri.WorldImagery) %>% 
  addLegend(colors = "red", labels = "Samples", title = "GEDI Level1B")
```


Get GEDI Full-waveform (GEDI Level1B)
```{r}
# Extracting GEDI full-waveform for a giving shotnumber
wf <- getLevel1BWF(gedilevel1b, shot_number = "33250017100000001")

par(mfrow = c(2, 1), mar = c(4, 4, 1, 1), cex.axis = 1.5)

plot(wf, relative=FALSE, polygon=TRUE, type="l", lwd=2, col="forestgreen",
     xlab = "Waveform Amplitude", ylab = "Elevation (m)")
grid()
plot(wf, relative=TRUE, polygon=FALSE, type="l", lwd=2, col="forestgreen",
     xlab="Waveform Amplitude (%)", ylab="Elevation (m)")
grid()
```


Get GEDI Elevation and Height Metrics
```{r}
# Get GEDI Elevation and Height Metrics
level2AM <- getLevel2AM(gedilevel2a)
head(level2AM[, c("beam", "shot_number", "elev_highestreturn", "elev_lowestmode", "rh100")])


# Converting shot_number as "integer64" to "character"
level2AM$shot_number <- paste0(level2AM$shot_number)

# Converting Elevation and Height Metrics as data.table to SpatialPointsDataFrame
level2AM$lon_lowestmode[is.na(level2AM$lon_lowestmode)] <- 0
level2AM$lat_lowestmode[is.na(level2AM$lat_lowestmode)] <- 0
level2AM_spdf <- SpatialPointsDataFrame(cbind(level2AM$lon_lowestmode, level2AM$lat_lowestmode),
                                        data=level2AM)

shot_number = "33250017100000001"

png("fig8.png", width = 8, height = 6, units = 'in', res = 300)
plotWFMetrics(gedilevel1b, gedilevel2a, shot_number, rh = c(25, 50, 75, 90))
dev.off()
```


Get GEDI Vegetation Biophysical Variables (GEDI Level2B)
```{r}
level2BVPM <- getLevel2BVPM(gedilevel2b)
head(level2BVPM[, c("beam", "shot_number", "pai", "fhd_normal", "omega", "pgap_theta", "cover")])



# Converting shot_number as "integer64" to "character"
level2BVPM$shot_number <- paste0(level2BVPM$shot_number)

# Converting GEDI Vegetation Profile Biophysical Variables as data.table to SpatialPointsDataFrame

level2BVPM$longitude_lastbin[is.na(level2BVPM$longitude_lastbin)] <- 0
level2BVPM$latitude_lastbin[is.na(level2BVPM$latitude_lastbin)] <- 0


level2BVPM_spdf <- SpatialPointsDataFrame(cbind(level2BVPM$longitude_lastbin, level2BVPM$latitude_lastbin), data = level2BVPM)
```


get GEDI plant area index and plant volume density
```{r}
level2BPAIProfile <- getLevel2BPAIProfile(gedilevel2b)
level2BPAVDProfile <- getLevel2BPAVDProfile(gedilevel2b)

# Converting shot_number as "integer64" to "character"
level2BPAIProfile$shot_number <- paste0(level2BPAIProfile$shot_number)
level2BPAVDProfile$shot_number <- paste0(level2BPAVDProfile$shot_number)

# Converting PAI and PAVD Profiles as data.table to SpatialPointsDataFrame
level2BPAIProfile$lon_lowestmode[is.na(level2BPAIProfile$lon_lowestmode)] <- 0
level2BPAIProfile$lat_lowestmode[is.na(level2BPAIProfile$lat_lowestmode)] <- 0

level2BPAIProfile_spdf <- SpatialPointsDataFrame(cbind(level2BPAIProfile$lon_lowestmode, level2BPAIProfile$lat_lowestmode),
data = level2BPAIProfile)



level2BPAVDProfile$lon_lowestmode[is.na(level2BPAVDProfile$lon_lowestmode)] <- 0
level2BPAVDProfile$lat_lowestmode[is.na(level2BPAVDProfile$lat_lowestmode)] <- 0

level2BPAVDProfile_spdf <- SpatialPointsDataFrame(cbind(level2BPAVDProfile$lon_lowestmode, level2BPAVDProfile$lat_lowestmode),
                                               data = level2BPAVDProfile)

# Exporting PAI and PAVD Profiles as ESRI Shapefile
# raster::shapefile(level2BPAIProfile_spdf,paste0(outdir,"\\GEDI02_B_2019108080338_O01964_T05337_02_001_01_sub_PAIProfile"))
# raster::shapefile(level2BPAVDProfile_spdf,paste0(outdir,"\\GEDI02_B_2019108080338_O01964_T05337_02_001_01_sub_PAVDProfile"))

```


Plot PAI and PAVD
```{r}
#specify GEDI beam
beam = "BEAM0101"

# Plot Level2B PAI Profile
gPAIprofile <- plotPAIProfile(level2BPAIProfile, beam = beam, elev = TRUE)

# Plot Level2B PAVD Profile
gPAVDprofile <- plotPAVDProfile(level2BPAVDProfile, beam = beam, elev = TRUE)
```

Clip all levels with boundary box
```{r}
level1bGeo_clip_bb <- clipLevel1BGeo(level1bGeo, xmin, xmax, ymin, ymax)
level2AM_clip_bb <- clipLevel2AM(level2AM, xmin, xmax, ymin, ymax)
level2BVPM_clip_bb <- clipLevel2BVPM(level2BVPM, xmin, xmax, ymin, ymax)
level1BPAIProfile_clip_bb <- clipLevel2BPAIProfile(level2BPAIProfile, xmin, xmax, ymin, ymax)
level2BPAVDProfile_clip_bb <- clipLevel2BPAVDProfile(level2BPAVDProfile, xmin, xmax, ymin, ymax)
```

Clip all levels with geometry
```{r}
#Need to run these to get descriptive statistics
level1bGeo_clip_gb <- clipLevel1BGeoGeometry(level1bGeo, study_Area, split_by = split_by)
level2AM_clip_gb <- clipLevel2AMGeometry(level2AM, polygon_spdf, split_by = split_by)
level2BVPM_clip_gb <- clipLevel2BVPMGeometry(level2BVPM, polygon_spdf, split_by = split_by)
level1BPAIProfile_clip_gb <- clipLevel2BPAIProfileGeometry(level2BPAIProfile, polygon_spdf, split_by = split_by)
level2BPAVDProfile_clip_gb <- clipLevel2BPAVDProfileGeometry(level2BPAVDProfile, polygon_spdf, split_by = split_by)
```


Display clipped level 12A with leaflet
```{r}
## View GEDI clipped data by bbox
library(leaflet)

m1 <- leaflet() %>%
  addCircleMarkers(level2AM_clip_bb$lon_lowestmode,
                   level2AM_clip_bb$lat_lowestmode,
                   radius = 1,
                   opacity = 1,
                   color = "green")  %>%
  addScaleBar(options = list(imperial = FALSE)) %>%
  addProviderTiles(providers$Esri.WorldImagery)  %>%
  addLegend(colors = c("green"), labels = c("Clip bbox"), title = "GEDI Level2A") 
```

Compute descriptive statistics of GEDI Level2A and Level2B data
```{r}
# Define your own function
mySetOfMetrics = function(x)
{
metrics = list(
    min = min(x), # Min of x
    max = max(x), # Max of x
    mean = mean(x), # Mean of x
    sd = sd(x)# Sd of x
  )
  return(metrics)
}

# Computing the maximum of RH100 stratified by polygon
rh100max_st <- polyStatsLevel2AM(level2AM_clip_bb,func = max(rh100))
head(rh100max_st)
```


Compute Grids with descriptive statistics of GEDI-derived Canopy Cover and Vertical Profile Metrics (Level2B)
```{r}
library(rasterVis)
library(viridis)
# Computing a serie of statistics of Total Plant Area Index
level2BVPM$pai[level2BVPM$pai == -9999] <- NA # assing NA to -9999
pai_metrics <- gridStatsLevel2BVPM(level2BVPM = level2BVPM, func = mySetOfMetrics(pai), res = 0.005)

# View maps
pai_maps <- levelplot(pai_metrics,
                    layout = c(1, 4),
                    margin = FALSE,
                    xlab = "Longitude (degree)", ylab = "Latitude (degree)",
                    colorkey = list(
                      space = 'right',
                      labels = list(at = seq(0, 1.5, 0.2), font = 4),
                      axis.line = list(col = 'black'),
                      width = 1),
                    par.settings = list(
                      strip.border = list(col = 'gray'),
                      strip.background = list(col = 'gray'),
                      axis.line = list(col = 'gray')
                    ),
                    scales = list(draw = TRUE),
                    col.regions = viridis,
                    at = seq(0, 1.5, len = 101),
                    names.attr = c("PAI min", "PAI max", "PAI mean", "PAI sd"))
```

