---
title: "accessingGEDI"
author: "Sam Watson"
date: "April 8, 2020"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{accessingGEDI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

load necessary packages
```{r}
library(gediamazonia)
library(rGEDI)
library(sf)
library(sp)
library(leaflet)
library(rasterVis)
library(viridis)
library(dplyr)
```

load AOI data and find GEDI data
```{r}
data(studyarea)
aoi <- studyarea$aoi
st_bbox(aoi)

#Create shapefiles from rda file
buffer <- studyarea$buffer
cacao_zone <- studyarea$cacao_zone
aoi <- studyarea$aoi
park <- studyarea$np


#create transect from area of interest, buffer, and cacao zone
aoi_buffer <- st_intersection(aoi, buffer)
transect <- st_intersection(aoi_buffer, cacao_zone)


study_Area <- st_read(dir("F:\\R_Project", pattern = "bb_points.shp", full.names = TRUE))
st_crs(study_Area) <- st_crs(aoi)

#bounding box coordinates
ul_lat <- -7.999644
lr_lat <- -9.437084
ul_lon <- -74.099156
lr_lon <- -76.157869

#Transect AOI coordinates
xmin <- -76.157869
xmax <- -74.099156
ymin <- -9.437084
ymax <- -7.999644

#Data Range
daterange=c("2019-08-01", "2019-08-15")
#Get path to GEDI

gLevel1B <- gedifinder(product = "GEDI01_B", ul_lat, ul_lon, lr_lat, lr_lon, version = "001", daterange = daterange)
gLevel2A <- gedifinder(product = "GEDI02_A", ul_lat, ul_lon, lr_lat, lr_lon, version = "001", daterange = daterange)
# gLevel2B <- gedifinder(product = "GEDI02_B", ul_lat, ul_lon, lr_lat, lr_lon, version = "001", daterange = daterange)
```

Downloading the data
```{r}
outdir = "F:\\R_Project\\July_2019"
outdir1 = "F:\\R_Project\\level1B"
outdir2 = "F:\\R_Project\\level2A"
outdir3 = "F:\\R_Project\\level2B"

# gediDownload(filepath = gLevel1B, outdir = outdir1)
# gediDownload(filepath = gLevel2A, outdir = outdir2)
# gediDownload(filepath = gLevel2B, outdir = outdir3)
```

reading GEDI data
```{r}
# Reading GEDI data
gedilevel1b <- readLevel1B(level1Bpath = paste0(outdir1, "\\GEDI01_B_2019216034003_O03638_T04321_02_003_01.h5"))
gedilevel2a <- readLevel2A(level2Apath = paste0(outdir2, "\\GEDI02_A_2019216034003_O03638_T04321_02_001_01.h5"))
gedilevel2b <- readLevel2B(level2Bpath = paste0(outdir3, "\\GEDI02_B_2019216034003_O03638_T04321_02_001_01.h5"))
```

GEDI pulse location level 1B
```{r}
#get geogprahic information from gedi data and extracting elevation
level1bGeo <- getLevel1BGeo(level1b = gedilevel1b, select = c("elevation_bin0"))
head(level1bGeo)


# Converting shot_number as "integer64" to "character"
level1bGeo$shot_number <- paste0(level1bGeo$shot_number)

# Converting level1bGeo as data.table to SpatialPointsDataFrame
library(sp)
level1bGeo$longitude_bin0[is.na(level1bGeo$longitude_bin0)] <- 0
level1bGeo$latitude_bin0[is.na(level1bGeo$latitude_bin0)] <- 0

#clipping level 1B GEDI to AOI
level1bGeo_clip_bb <- clipLevel1BGeo(level1bGeo, xmin, xmax, ymin, ymax)

level1bGeo_spdf <- SpatialPointsDataFrame(cbind(level1bGeo$longitude_bin0, level1bGeo$latitude_bin0),
                                        data = level1bGeo)

```


Get GEDI Full-waveform (GEDI Level1B)
```{r}
# Extracting GEDI full-waveform for a giving shotnumber
wf <- getLevel1BWF(gedilevel1b, shot_number = "36380009200000001")

par(mfrow = c(2, 1), mar = c(4, 4, 1, 1), cex.axis = 1.5)

plot(wf, relative=FALSE, polygon=TRUE, type="l", lwd=2, col="forestgreen",
     xlab = "Waveform Amplitude", ylab = "Elevation (m)")
grid()
plot(wf, relative=TRUE, polygon=FALSE, type="l", lwd=2, col="forestgreen",
     xlab="Waveform Amplitude (%)", ylab="Elevation (m)")
grid()
```


Get GEDI Elevation and Height Metrics
```{r}
# Get GEDI Elevation and Height Metrics
level2AM <- getLevel2AM(gedilevel2a)
head(level2AM[, c("beam", "shot_number", "elev_highestreturn", "elev_lowestmode", "rh100")])


# Converting shot_number as "integer64" to "character"
level2AM$shot_number <- paste0(level2AM$shot_number)

# Converting Elevation and Height Metrics as data.table to SpatialPointsDataFrame
level2AM$lon_lowestmode[is.na(level2AM$lon_lowestmode)] <- 0
level2AM$lat_lowestmode[is.na(level2AM$lat_lowestmode)] <- 0

level2AM_spdf <- SpatialPointsDataFrame(cbind(level2AM$lon_lowestmode, level2AM$lat_lowestmode),
                                        data=level2AM)

shot_number = "36380009200000001"

png("fig8.png", width = 8, height = 6, units = 'in', res = 300)
plotWFMetrics(gedilevel1b, gedilevel2a, shot_number, rh = c(25, 50, 75, 90))
dev.off()
```


Get GEDI Vegetation Biophysical Variables (GEDI Level2B)
```{r}
level2BVPM <- getLevel2BVPM(gedilevel2b)
head(level2BVPM[, c("beam", "shot_number", "pai", "fhd_normal", "omega", "pgap_theta", "cover")])



# Converting shot_number as "integer64" to "character"
level2BVPM$shot_number <- paste0(level2BVPM$shot_number)

# Converting GEDI Vegetation Profile Biophysical Variables as data.table to SpatialPointsDataFrame

level2BVPM$longitude_lastbin[is.na(level2BVPM$longitude_lastbin)] <- 0
level2BVPM$latitude_lastbin[is.na(level2BVPM$latitude_lastbin)] <- 0


level2BVPM_spdf <- SpatialPointsDataFrame(cbind(level2BVPM$longitude_lastbin, level2BVPM$latitude_lastbin), data = level2BVPM)
```


get GEDI plant area index and plant volume density
```{r}
level2BPAIProfile <- getLevel2BPAIProfile(gedilevel2b)
level2BPAVDProfile <- getLevel2BPAVDProfile(gedilevel2b)

# Converting shot_number as "integer64" to "character"
level2BPAIProfile$shot_number <- paste0(level2BPAIProfile$shot_number)
level2BPAVDProfile$shot_number <- paste0(level2BPAVDProfile$shot_number)

# Converting PAI and PAVD Profiles as data.table to SpatialPointsDataFrame
level2BPAIProfile$lon_lowestmode[is.na(level2BPAIProfile$lon_lowestmode)] <- 0
level2BPAIProfile$lat_lowestmode[is.na(level2BPAIProfile$lat_lowestmode)] <- 0

level2BPAIProfile_spdf <- SpatialPointsDataFrame(cbind(level2BPAIProfile$lon_lowestmode, level2BPAIProfile$lat_lowestmode),
data = level2BPAIProfile)



level2BPAVDProfile$lon_lowestmode[is.na(level2BPAVDProfile$lon_lowestmode)] <- 0
level2BPAVDProfile$lat_lowestmode[is.na(level2BPAVDProfile$lat_lowestmode)] <- 0

level2BPAVDProfile_spdf <- SpatialPointsDataFrame(cbind(level2BPAVDProfile$lon_lowestmode, level2BPAVDProfile$lat_lowestmode),
                                               data = level2BPAVDProfile)

```


Plot PAI and PAVD
```{r}
#specify GEDI beam
beam = "BEAM0101"

# Plot Level2B PAI Profile
gPAIprofile <- plotPAIProfile(level2BPAIProfile, beam = beam, elev = TRUE)

# Plot Level2B PAVD Profile
gPAVDprofile <- plotPAVDProfile(level2BPAVDProfile, beam = beam, elev = TRUE)
```

Clip all levels with boundary box
```{r}
level1bGeo_clip_bb <- clipLevel1BGeo(level1bGeo, xmin, xmax, ymin, ymax)
level2AM_clip_bb <- clipLevel2AM(level2AM, xmin, xmax, ymin, ymax)
level2BVPM_clip_bb <- clipLevel2BVPM(level2BVPM, xmin, xmax, ymin, ymax)
level1BPAIProfile_clip_bb <- clipLevel2BPAIProfile(level2BPAIProfile, xmin, xmax, ymin, ymax)

#create spatial dataframes from clipped levels
clipped_PAI_spdf <- SpatialPointsDataFrame(cbind(level1BPAIProfile_clip_bb$lon_lowestmode, level1BPAIProfile_clip_bb$lat_lowestmode),
                                               data = level1BPAIProfile_clip_bb)

clipped_PAVD_spdf <- clipLevel2BPAVDProfile(level2BPAVDProfile, xmin, xmax, ymin, ymax)
# Exporting PAI and PAVD Profiles as ESRI Shapefile
raster::shapefile(clipped_PAI_spdf,paste0(outdir,"\\PAIProfile_CLIP")) 




raster::shapefile(clipped_PAVD_spdf,paste0(outdir,"\\PAVDProfile_CLIP"))
```

Clip all levels with geometry
```{r}
#Need to run these to get descriptive statistics
library(rgdal)
polygon_spdf <- readOGR("C:\\Sam_Scratch\\Spatial_R\\gediamazonia\\merged.shp")
split_by = "id"

level1bGeo_clip_gb <- clipLevel1BGeoGeometry(level1bGeo, polygon_spdf, split_by = split_by)
level2AM_clip_gb <- clipLevel2AMGeometry(level2AM, polygon_spdf, split_by = split_by)
level2BVPM_clip_gb <- clipLevel2BVPMGeometry(level2BVPM, polygon_spdf, split_by = split_by)
level1BPAIProfile_clip_gb <- clipLevel2BPAIProfileGeometry(level2BPAIProfile, polygon_spdf, split_by = split_by)
level2BPAVDProfile_clip_gb <- clipLevel2BPAVDProfileGeometry(level2BPAVDProfile, polygon_spdf, split_by = split_by)
```


Display clipped level 12A with leaflet
```{r}
# View GEDI clipped data by bbox

leaflet() %>%
  addCircleMarkers(level2BVPM_clip_gb$lon_lowestmode,
                   level2BVPM_clip_gb$lat_lowestmode,
                   radius = 1,
                   opacity = 1,
                   color = "green")  %>%
  addScaleBar(options = list(imperial = FALSE)) %>%
  addProviderTiles(providers$Esri.WorldImagery)  %>%
  addLegend(colors = c("green"), labels = c("Clip bbox"), title = "GEDI Level2A")
```

Compute descriptive statistics of GEDI Level2A and Level2B data
```{r}
# Define your own function
mySetOfMetrics = function(x)
{
metrics = list(
    min = min(x), # Min of x
    max = max(x), # Max of x
    mean = mean(x), # Mean of x
    sd = sd(x)# Sd of x
  )
  return(metrics)
}

# Computing the maximum of RH100 stratified by polygon
rh100max_st <- polyStatsLevel2AM(level2AM_clip_bb,func = max(rh100))
head(rh100max_st)

# Computing a serie statistics for GEDI metrics stratified by polygon
rh100metrics_st <- polyStatsLevel2AM(level2AM_clip_gb,func = mySetOfMetrics(rh100),
id = "poly_id")
head(rh100metrics_st)


# Computing the max of the Total Plant Area Index
pai_max <- polyStatsLevel2BVPM(level2BVPM_clip_bb,func = max(pai), id = NULL)
pai_max


# Computing a series of statistics of Canopy Cover stratified by polygon
PAI_NA <- level2BVPM_clip_gb$pai[level2BVPM_clip_gb$pai == -9999] <- NA
pai_metrics_st <- polyStatsLevel2BVPM(level2BVPM_clip_gb,func = mySetOfMetrics(pai),
id = "poly_id")
head(pai_metrics_st)

         
```


Compute Grids with descriptive statistics of GEDI-derived Canopy Cover (Level2B)
```{r}
# Computing a series of statistics of Total Plant Area Index and plotting in leaflet
 # assing NA to -9999
pai_metrics <- gridStatsLevel2BVPM(level2BVPM = level2BVPM_clip_gb, func = mySetOfMetrics(pai), res = 0.005)

#create mean PAI layer from rasterbrick, project, and mask
meanPAI <- pai_metrics$mean 
crs(meanPAI) <- "+proj=longlat +datum=WGS84 +no_defs"

meanPAI_CaocaoMask <- meanPAI %>% mask(x = ., mask = cacao_zone) %>% as.data.frame(.) %>% mutate(zone = "Cacao")

meanPAI_ParkMask <- meanPAI %>% mask(x = ., mask = park) %>% as.data.frame(.) %>% mutate(zone = "Park")

meanPAI_BufferMask <- meanPAI %>% mask(x = ., mask = buffer) %>% as.data.frame(.) %>% mutate(zone = "Buffer")

pai_zones <- rbind(meanPAI_ParkMask, meanPAI_CaocaoMask, meanPAI_BufferMask)
pai_plot <- pai_zones %>% ggplot() + geom_boxplot(aes(x = zone, y = mean)) + xlab(NULL)
ggsave("pai_zones.png", pai_plot, width = 2, height = 1)


pal <- colorNumeric(c("#0C2C84", "#41B6C4", "#FFFFCC"), values(meanPAI_Mask),
  na.color = "transparent")

leaflet() %>% addTiles() %>% addRasterImage(meanPAI, opacity = 0.8) %>%
  addPolygons(data = park, noClip = T,
              weight = 4,
              dashArray = "5, 1",
              color = "black",
              fillOpacity = .01,
              smoothFactor = 0) %>% 
   addPolygons(data = cacao_zone, noClip = T,
              weight = 4,
              dashArray = "5, 1",
              color = "red",
              fillOpacity = .01,
              smoothFactor = 0) %>% 
   addPolygons(data = buffer, noClip = T,
              weight = 4,
              dashArray = "5, 1",
              color = "blue",
              fillOpacity = .01,
              smoothFactor = 0)

```




Get Relative Height Metrics and plot
```{r}
# Computing a serie of statistics of GEDI RH100 metric
rh100metrics <- gridStatsLevel2AM(level2AM = level2AM_clip_gb, func = mySetOfMetrics(rh100), res = 0.005)

meanRH100 <- rh100metrics$mean 
crs(meanRH100) <- "+proj=longlat +datum=WGS84 +no_defs"

meanRH100_CaocaoMask <- meanRH100 %>% mask(x = ., mask = cacao_zone) %>% as.data.frame(.) %>% mutate(zone = "Cacao", level = "rh100")

meanRH100_ParkMask <- meanRH100 %>% mask(x = ., mask = park) %>% as.data.frame(.) %>% mutate(zone = "Park", level = "rh100")

meanRH100_BufferMask <- meanRH100 %>% mask(x = ., mask = buffer) %>% as.data.frame(.) %>% mutate(zone = "Buffer", level = "rh100")

rh100levels <- rbind(meanRH100_CaocaoMask, meanRH100_ParkMask,meanRH100_BufferMask)

# Computing a serie of statistics of GEDI RH75 metric
rh75metrics <- gridStatsLevel2AM(level2AM = level2AM_clip_gb, func = mySetOfMetrics(rh75), res = 0.005)

meanRH75 <- rh75metrics$mean 
crs(meanRH75) <- "+proj=longlat +datum=WGS84 +no_defs"

meanRH75_CaocaoMask <- meanRH75 %>% mask(x = ., mask = cacao_zone) %>% as.data.frame(.) %>% mutate(zone = "Cacao", level = "rh75")

meanRH75_ParkMask <- meanRH75 %>% mask(x = ., mask = park) %>% as.data.frame(.) %>% mutate(zone = "Park", level = "rh75")

meanRH75_BufferMask <- meanRH75 %>% mask(x = ., mask = buffer) %>% as.data.frame(.) %>% mutate(zone = "Buffer", level = "rh75")

rh75levels <- rbind(meanRH75_CaocaoMask, meanRH75_ParkMask,meanRH75_BufferMask)

# Computing a serie of statistics of GEDI RH50 metric
rh50metrics <- gridStatsLevel2AM(level2AM = level2AM_clip_gb, func = mySetOfMetrics(rh50), res = 0.005)

meanRH50 <- rh50metrics$mean 
crs(meanRH75) <- "+proj=longlat +datum=WGS84 +no_defs"

meanRH50_CaocaoMask <- meanRH50 %>% mask(x = ., mask = cacao_zone) %>% as.data.frame(.) %>% mutate(zone = "Cacao", level = "rh50")

meanRH50_ParkMask <- meanRH50 %>% mask(x = ., mask = park) %>% as.data.frame(.) %>% mutate(zone = "Park", level = "rh50")

meanRH50_BufferMask <- meanRH50 %>% mask(x = ., mask = buffer) %>% as.data.frame(.) %>% mutate(zone = "Buffer", level = "rh50")

rh50levels <- rbind(meanRH50_CaocaoMask, meanRH50_ParkMask,meanRH50_BufferMask)

# Computing a serie of statistics of GEDI RH25 metric
rh25metrics <- gridStatsLevel2AM(level2AM = level2AM_clip_gb, func = mySetOfMetrics(rh25), res = 0.005)

meanRH25 <- rh25metrics$mean 
crs(meanRH25) <- "+proj=longlat +datum=WGS84 +no_defs"

meanRH25_CaocaoMask <- meanRH25 %>% mask(x = ., mask = cacao_zone) %>% as.data.frame(.) %>% mutate(zone = "Cacao", level = "rh25")

meanRH25_ParkMask <- meanRH25 %>% mask(x = ., mask = park) %>% as.data.frame(.) %>% mutate(zone = "Park", level = "rh25")

meanRH25_BufferMask <- meanRH25 %>% mask(x = ., mask = buffer) %>% as.data.frame(.) %>% mutate(zone = "Buffer", level = "rh25")

rh25levels <- rbind(meanRH25_CaocaoMask, meanRH25_ParkMask,meanRH25_BufferMask)


#Bind all rows
rhlevels <- rbind(rh100levels, rh75levels, rh50levels, rh25levels)

#plot rh levels in boxplot for zones
rhplot <- rhlevels %>% 
  ggplot() + geom_boxplot(aes(x = level, y = mean, fill = zone)) + 
  xlab(NULL) + scale_fill_manual(values = c("Dark Green", "white", "light blue")) + ggtitle("Grid Mean Relative Height Metrics")
ggsave("rhplot.png", rhplot)


```

