---
title: "Visualizing GEDI Data"
author: "Amanda Payton, Sam Watson, Wenqu Chen"
date: "May 2, 2020"
output: html_document
---

# Overview

### This rmarkdown uses the clipped shapefiles from the GEDI data vignette and 
### visualizes the data through maps and box plots.

# Setup

## Package Installation and library loading

```{r}
#install.packages("rGEDI")
library(rGEDI)
#install.packages("geospaar")
library(geospaar)
library(ggplot2)
```

## Set Folder paths

```{r}
#check working directory
getwd() 
#If needed, set your working directory to the gediamazonia project folder

data_fold = paste0(getwd(), "/data")
working_fold = paste0(getwd(), "/working")
```

## Read in GEDI Shapefiles and Study Area Data

```{r}
data("studyarea")
data("roads")
data('gedi_ss')

#Data clipped to AOI
level1b <- st_read(dsn = paste0(data_fold, "/aoi_GEDI_1B.shp"))
head(level1b)

level2a <- st_read(dsn = paste0(data_fold, "/aoi_GEDI_2A.shp"))
head(level2a)

level2b <- st_read(dsn = paste0(data_fold, "/aoi_GEDI_2B.shp"))
head(level2b)

level2bPAI <- st_read(dsn = paste0(data_fold, "/aoi_GEDI_2B_PAI.shp"))
head(level2bPAI)

level2bPAVD <- st_read(dsn = "/aoi_GEDI_2B_PAVD.shp")
head(level2bPAVD)

#Data clipped to entire study region
level2a_SA <- st_read(dsn = paste0(data_fold, "/SA_GEDI_2A.shp"))
head(level2a_SA)

level2b_SA <- st_read(dsn = paste0(data_fold, "/SA_GEDI_2B.shp"))
head(level2b_SA)
```

# Visualizations

## Map Study Area

```{r}
#Create map of Study Area
plot_SA <- ggplot() + geom_sf(data = studyarea$np, fill = "dark green") + 
  geom_sf(data = studyarea$buffer, color = "grey", fill = "light blue") + 
  geom_sf(data = studyarea$cacao_zone, color = "brown", 
          fill = alpha("tan", .2)) +   
  geom_sf(data = roads, color = "black") +
  theme_minimal() +
  ggtitle("Cordillera Azul National Park") + 
     theme(plot.title = element_text(face = "bold"))
#Display map
plot_SA

#Export to Working Folder
ggsave(plot_SA, filename = paste0(working_fold, "/study_area.png"), height = 4,
       width = 5, dpi = 300)
```

## Map GEDI Variables

```{r}
#Create map of metrics
rh_points <- ggplot() + geom_sf(data = studyarea$np, color = "grey", 
                                fill = "grey70") + 
  geom_sf(data = studyarea$buffer, color = "grey", fill = "grey80") + 
  geom_sf(data = studyarea$cacao_zone, color = "grey", fill = "grey90") +   
  geom_sf(data = studyarea$aoi, color = "grey") + 
  geom_sf(data = roads, color = "grey") +
  theme_minimal() +
  #geom_sf(data = level1b, aes(color = elvtn_0)) + #level 1B Elevation data
  #geom_sf(data = level2a, aes(color = rh25)) + #level 2A relative height 25
  #geom_sf(data = level2a, aes(color = rh50)) + #level 2A relative height 50
  #geom_sf(data = level2a, aes(color = rh75)) + #level 2A relative height 75
  #geom_sf(data = level2a, aes(color = rh100)) + #level 2A relative height 100
  #geom_sf(data = level2b, aes(color = pai)) + #level 2B plant area index
  geom_sf(data = level2a_SA, aes(color = rh100)) + #level 2A rh 100 for study area
  #geom_sf(data = level2b_SA, aes(color = pai)) + #level 2A pai for study area
  ggtitle("Relative Height lvl 100 Distribution")
#Display map
rh_points

#Export to Working Folder
ggsave(paste0(working_fold, "/rh_points.png"), rh_points, width = 5, height = 4, 
       dpi = 300)
```

## Create Boxplots - Relative Height and Zones

### Preprocessing

```{r}
#Intersect level 2A data shapefile with zones 
cacao_intersect <- st_intersection(x = studyarea$cacao_zone, y = level2a_SA)
park_intersect <- st_intersection(x = studyarea$np, y = level2a_SA)
buffer_intersect <- st_intersection(x = studyarea$buffer, y = level2a_SA)

#Create Relative Height 25 Variable
cacao_RH25 <- cacao_intersect$rh25 %>% as_data_frame(.) %>% 
  mutate(zone = "Cacao", level = "rh25")
park_RH25 <- park_intersect$rh25 %>% as_data_frame(.) %>% 
  mutate(zone = "Park", level = "rh25")
buffer_RH25 <- buffer_intersect$rh25 %>% as_data_frame(.) %>% 
  mutate(zone = "Buffer", level = "rh25")

#Create Relative Height 50 Variable
cacao_RH50 <- cacao_intersect$rh50 %>% as_data_frame(.) %>% 
  mutate(zone = "Cacao", level = "rh50")
park_RH50 <- park_intersect$rh50 %>% as_data_frame(.) %>% 
  mutate(zone = "Park", level = "rh50")
buffer_RH50 <- buffer_intersect$rh50 %>% as_data_frame(.) %>% 
  mutate(zone = "Buffer", level = "rh50")

#Create Relative Height 75 Variable
cacao_RH75 <- cacao_intersect$rh75 %>% as_data_frame(.) %>% 
  mutate(zone = "Cacao", level = "rh75")
park_RH75 <- park_intersect$rh75 %>% as_data_frame(.) %>% 
  mutate(zone = "Park", level = "rh75")
buffer_RH75 <- buffer_intersect$rh75 %>% as_data_frame(.) %>% 
  mutate(zone = "Buffer", level = "rh75")

#Create Relative Height 100 Variable
cacao_RH100 <- cacao_intersect$rh100 %>% as_data_frame(.) %>% 
  mutate(zone = "Cacao", level = "rh100")
park_RH100 <- park_intersect$rh100 %>% as_data_frame(.) %>% 
  mutate(zone = "Park", level = "rh100")
buffer_RH100 <- buffer_intersect$rh100 %>% as_data_frame(.) %>% 
  mutate(zone = "Buffer", level = "rh100")

#bind all RH levels
rh_bound <- rbind(cacao_RH25, park_RH25, buffer_RH25, cacao_RH50, 
                  park_RH50, buffer_RH50, cacao_RH75, park_RH75, 
                  buffer_RH75, cacao_RH100, park_RH100, buffer_RH100)

#Display mean rh value of each level and zone
rh_bound %>% group_by(zone, level) %>% summarise_all(funs(mean))
```

### Plot

```{r}
#plot RH levels
rhplot <- rh_bound %>% ggplot() + 
  geom_boxplot(aes(x = level, y = value, fill = zone)) + xlab(NULL) + 
  scale_fill_manual(values = c("light blue", "tan", "dark green")) + 
  ggtitle("Relative Height Metrics Between Geometries")

#save plot
ggsave(rhplot, filename = paste0(working_fold, "rh_zone_plot.png", width = 5, 
                                 height = 3))
```

## Create Boxplots - Plant Area Index and Zones

### Preprocessing

```{r}
#Remove -9999 from PAI metric
level2b_SA[level2b_SA$pai == -9999] <- NA

#clip clipped shapefile to different zones
cacao_PAI <- st_intersection(x = cacao_zone, y = level2b_SA)
park_PAI <- st_intersection(x = park, y = level2b_SA)
buffer_PAI <- st_intersection(x = buffer, y = level2b_SA)

#make PAI dataframes for each zone
cacao_PAI <- cacao_PAI$pai %>% as_data_frame(.) %>% mutate(zone = "Cacao")
park_PAI <- park_PAI$pai %>% as_data_frame(.) %>% mutate(zone = "Park")
buffer_PAI <- buffer_PAI$pai %>% as_data_frame(.) %>% mutate(zone = "Buffer")

#bind PAI zones
pai_bound <- rbind(cacao_PAI, park_PAI, buffer_PAI)

#Display mean PAI value of each zone
pai_bound %>% group_by(zone) %>% summarise_all(funs(mean))
```

### Plot

```{r}
#set color pallette
clrs <- c("light blue", "tan", "dark green")

#plot PAI
paiplot <- ggplot() + 
  geom_boxplot(aes(x = zone, y = value, fill = zone)) + xlab(NULL) + 
  ggtitle("Plant Area Index Metrics by Zone") + scale_fill_manual(values = clrs)

#save plot
ggsave(paiplot, filename = paste0(working_fold, "pai_zones_plot.png"), 
       width = 5, height = 3)
```

## Create Boxplots - Relative Height and Road Distance

### Preprocessing

```{r}
#Create dataframes for each rh level
rh25 <- gedi_ss$rh25 %>% as_data_frame(.) %>% mutate(level = "rh25")
rh50 <- gedi_ss$rh50 %>% as_data_frame(.) %>% mutate(level = "rh50")
rh75 <- gedi_ss$rh75 %>% as_data_frame(.) %>% mutate(level = "rh75")
rh100 <- gedi_ss$rh100 %>% as_data_frame(.) %>% mutate(level = "rh100")
road_class <- gedi_ss$distclass %>% as_data_frame(.) %>% 
  mutate(distances = value)

#bind all of the rh levels and road class
allrhlevels <- rbind(rh25, rh50, rh75, rh100) #bind rh level
gediss_bind <- cbind(road_class, allrhlevels) #add road class
road_rh_levels <- gediss_bind[-c(1)] #remove extra value column

#Display mean PAI value of each zone
road_rh_levels %>% group_by(distances, level) %>% summarise_all(funs(mean))
```

### Plot

```{r}
roads_rhplot <- road_rh_levels %>% ggplot() + 
  geom_boxplot(aes(x = level, y = value, fill = distances)) + xlab(NULL) + 
  scale_fill_manual(values = c("#f0f9e8", "#ccebc5", "#a8ddb5", "#7bccc4", 
                               "#43a2ca", "#0868ac")) + 
  ggtitle("Relative Height Metrics Dist. from Roads") + 
  labs(fill = "Distance Rank")

ggsave(roads_rhplot, filename = paste0(working_fold, "rh_roads_plot.png"), 
       width = 5, height = 3)
```
